---
title: "Final Project 134"
author: "Fion, Jessica, Dulce, Yikai"
date: "2025-02-27"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
editor_options: 
  markdown: 
    wrap: sentence
---

# **Introduction to the Recommendation System Project**

## Project Overview

In this project, our group aims to develop a recommendation system using an Amazon dataset to enhance user experience by providing personalized product recommendations.
With the vast amount of data available in e-commerce, recommendation systems play a crucial role in helping users discover relevant products efficiently.
By leveraging machine learning techniques, we aim to build a model that predicts user preferences based on historical interactions and product attributes.

## Data Description

## Methodology

## Project Road Map

## Exploring Our Data

## Loading Packages

```{r}
library(tidyverse)
library(dplyr)
```

## Loading in the Dataset

```{r}
# load the data set
amazon <- read.csv("data/amazon.csv")
```

## Exploring Raw Data

```{r}
summary(amazon)
colSums(is.na(amazon))
```

## Tidying the Data

```{r}
# Mutate character variables into numeric
amazon$discounted_price <- as.numeric(gsub("[^0-9.]", "", amazon$discounted_price))
amazon$actual_price <- as.numeric(gsub("[^0-9.]", "", amazon$actual_price))
amazon$discount_percentage <- as.numeric(gsub("[^0-9]", "", amazon$discount_percentage))
amazon$rating_count <- as.numeric(gsub("[^0-9.]", "", amazon$rating_count))

# Find non-numeric rating values
print(amazon$rating[grepl("[^0-9.]", amazon$rating)])

amazon <- amazon %>% 
  filter(grepl("^[0-9.]", rating)) %>% 
  filter(!is.na(rating_count)) %>% 
  mutate(rating = as.numeric(rating))

# Confirm no missing
colSums(is.na(amazon))
```

posssible pca, nlp usage

```{r}
summary(amazon)
```

## Visual EDA

# 
